Summer 2025 ToDo

### 2025-11-10 Relations/Paper/Introduction.tex
Preparing to submit the paper

This broader goal also drives the main design choices in our approach.
Since rewriting theory is concerned with a fine-grained analysis of computation,
the most natural vehicle for formalizing it is type theory based on the
Curry--Howard isomorphism, which Agda implements.  Proofs in this language
are automatically effective, and implicitly carry the code that implements
a transformation from the hypotheses to the conclusion of the claim.  For example,
a type-checked proof that a given TRS is confluent automatically renders a function that
computes a common reduct between any two reducts of a given term.

>>>>

This broader goal also drives the main design choices in our approach.
Since rewriting theory is concerned with a fine-grained analysis of computation,
the most natural vehicle for formalizing it is type theory based on the
Curry--Howard isomorphism, which Agda implements.  Proofs in this language
are automatically effective, and implicitly carry the code that implements
a transformation from the hypotheses to the conclusion of the claim.  For example,
a type-checked proof that a given TRS is Church--Rosser automatically renders a function that
computes a common reduct between any two convertible terms.

### 2025-11-10 Relations/Paper/Summer_2025_TODO.tex
Preparing to submit the paper

    \item (Review 3) Check out and include the following related work:

@@

\begin{itemize}
  \item Perhaps we should not create variations and multiple implications of WFind,
  but rather prove that it's equivalent to WFacc, but lands in a higher universe.
  Mention this once at the beginning of section 5 then proceed to analyze
  connections between WFacc and WFseq/WFmin.
\end{itemize}

\section*{July TODO}
\begin{todolist}
  \item Carry on refactoring Wellfounded.agda once fixed, breaking it up if necessary.
  \item To finish the paper: We should do a final walk through and make sure we haven't introduced errors with our edits. In particular, do all the counterexamples line up. Do all the references to different sections and pieces of code in Agda line up. 
  \item Do we remove all our local counterexamples in favour of global ones?
  \item Rework well-foundedness section 
  \item Rework conclusion 
  \item Rework introduction 
  \item Do we introduce a new section for discussing Agda? Talking about why we think there is an advantage to doing this work in Agda vs (say) Isabelle

@@

\section*{Research items}
\begin{todolist}
  \item (Research item) Go deeper in the related works and see which are most relevant. Make a list of works and give summaries to be used in future papers. * Look into the literature and related works. There should be an Agda formulation of rewrite systems (check agda standard library stuff on relations). 
  \item (Research item) Find a concrete application of generalized Newman's Lemma. * Finding a term rewrite system where it's eassier to see that it is weakly cr and that every term is SM (as opposed to being CR).  
  \item (Research item) Formalize a first order term rewriting system (I can't remember what the motivation for this was). Use TeReSe to create a list of rewrite systems.
  \item (Research item) Carry out research on rewriting counterexamples like our own so we can try and put them into some sort of context. Have our counerexamples been shown before?
  \item (Research item) Work on the missing implicaitons from the paper. So far we believe this is just CR $\to$ CP in the ARS section. - Will probably be quite a bit of work (low priority)
  \item (Research item) Do we want to reincorporate Knaster Tarski? - Not this paper 
  \item (Research item) How to go from WFseq to WFacc - possible that the Kleene Tree is a solution. - Not this paper

\section*{Agda standard library}
Some of the work we have explored is formalized in the Agda standard library.  
% https://agda.github.io/agda-stdlib/v1.7.3/Relation.Binary.Rewriting.html#2096
\begin{todolist}
  \item Write about the standard library formalization in our paper and talk about how ours differs. 
  \item Create an equivalences file showing that our formalization agrees with the ASL (specifically definitions such as confluet, wcr, nf, leading up to showing our formalizations of Newman's lemma agree).
  \item Investigate contributing work to the ASL. This may mean refactoring the work already done to conform with the ASL. Specific contributions would likely be: Generalized NL. Theorem's 122 and 123. Alternative definitions of Wellfounded, the implications for the hierarchy.
\end{todolist}

\section*{20th August todos}
\begin{todolist}
  \item [\done] (Secondary) Find consistency with NP and UN with respect to reduction and conversion. Should the default be by reduction for both. Should we be explicit with $\to$ and = in each case.  \sacomment{Main issue here is that we'll either introduce an inconsistency with our code or we'll have a hard to read symbol. We can do $\to$ as super or sub-script in Agda. If we're happy with $NP\to$ in agda but using a different script in the paper then I think we should go ahead and be explicit for both UN and NP as to whether it is WRT conversion or reduction. We already have this inconsistency in the code with $UN\to$} \apcomment{In the agda code let's not use arrow and have arrow implied as default but be explicit with '=' sign.}
  \item Carry on modifying the names of wfXXX- to wfXXXnotnot with prefix underscore in agda and in latex. \sacomment{Done with the exception of WFseq-. This needs a bit more thought. Similarly for the relation between WFcor and WFminCor.}
  \item [\done] Make a decision regarding WFcor. We use it to imply WFacc using AccCor. But otherwise it is sort of orphaned off. Does it help contribute to the overall story? \sacomment{We've now related it to WFmin by requiring that correductive predicates have double negation elimination. Is this too strong a requirement? Similarly we have linked WFacc to WFcor but this proof requires the same property as well as limiting ourselves to accessible predicates. Again, is this too much?}
  \item [\done] Get the classical hypothesis and relations organized and see how they're used in the implications (organize implications by classical requirement. In the implications file we can import the classical implications. Basically, refactor the code.) \sacomment{Done but could be further refactored.}
  \item Make a cheat sheet for the next meeting regarding to talk about coreductive. How it relates on the weak side and on the strong side. SEQ seq+- , how to present these. Try and clarify what the issues are here and what exactly are the differences between them in terms of definition and implication. Could one be removed and the overall story stay the same? \sacomment{Regarding seq+ (and seq+-) we currently only show that seq+ implies seq and seq+- implies seq- so could be removed without changing any of our other implictions.} 
\end{todolist}

\section*{September 15th todo}
\begin{todolist}
  \item (Minor) We want to have a sentence in our WF section making clear that we can consider local properties as well and maybe make an example from accessible to sequential. 
  \item Sentence explaining that wfacc is equivalent to wfind and therefore isn't included in the diagram (wfind).
\end{todolist}

\section*{September 26th TODO}
\begin{todolist}     
  \item Focus for next meeting on writing. 
  \item [\done] Are not not closed predicates over natural numbers dedcidable: If not is this a counterexample for WFminDNE.
\end{todolist}

\section*{October 10th TODO}
\begin{todolist}     
  \item Want to show that FB and weak decidability gives us accCor (think we may have shown this previously)
  \item Add 'is' to wfcor macros.
  \item make a decision on wfseq not not name. 
  \item Keep on cleaning the WF section in paper. 
  \item 
\end{todolist}

\section*{October 15th TODO / Oct 22nd}
\begin{todolist}     
  \item It's possible to have a term rewrite system with good properties like FB and the relation being decidable. But not satisfying accCor. That is something that can be added in a final touch of the paper. Either a remark or example, where a term rewrite system does not validate accCor. (In the lambda calculus, there are computable perpetual strategies. We think they're concerned with weak normalization. Not sure if it can be done with strong normalization.)
  \item Do a final read through of the information below the figure. Have we captured all the most important information / Are we repeating ourselves?
  \item [\done] Do we want to keep anything in the old/delete section of WF?
  \item Do we want to add anything about counter examples or positive examples? 
  \item [\done] We may want to use words rather than macros in certain locations
  \item [\done] Check through the comments in general in the section. 
  \item (Long term) Investigate: what we've done for natural numbers could be done for an arbitrary ARS (in the counter examples files). The sentence: ''The restriction to $\lnot\lnot$-closed subsets implies (and is equivalent to) 
every subset $P$ being merely \emph{weakly} decidable (satisfying $\lnot \lnot P(x) \lor \lnot P(X)$ 
for all $x$)'' needs further exploration. We could not say it/say it differently. 
\item [\done]  CHange WFseq not not to WFseq- in the macro. We are sticking with this for now. If we have time we can investigate the relation between the two.
\item [\done]  Make the diagram horizontal rather than vertical. Spoof it up and we can compare next week.
\item [\done]  Use of macros in WF section. Double check the right one is being used. Don't use the "is'' macro at all in this section. Be consistent on whether we are using macros. 
\item Does WFseq- with MP== imply WFseq? (W're thinking it will need rdec as well to give us the implication).
\item [\done] make the WF macro bold font?
\item add wfseq not not to the definitions?
\item [\done] Go through the bibliography links. Not all of them are currently working!
\item We want to start looking at examples. What we want to know is whether a first order rewrite system with some minimal properties implies some of our classical properties. E.g. lambda calculus satisfies some of our classical properties. As does first order term rewrite system. Follow up question: how big of an issue is it that accCor is not provable? Can we prove is isn't satisfied? 
\item [\done] Make a connection in the final WF section talking about how SN doesn't imply WN constructively. (we've started writing this sentence). 
\item [\done] We may need to add to the WFCounters file, if we have SN implies WN then every subset of the natural numbers is decidable.  
\item Make sure we have a proof of the implication from SN to WN given FB and decidability of a relation (to back up our claim in our introduction).
\item check TERESE and see first order term rewrite systems example which could be valid for showing classical and constructive formalization coincide. First order term rewrite systems are (always?) fb and R is dec. 
\item [\done] Change the macro for FB so that it is also boldface
\item Need to fix how we put urls in for bibliography. Currently not quite working. 
\item update wfcounters to show that asserting wfmin implies the excluded middle (and wfmindne the weak excluded middle).
\item [\done] In wfcounters we should have a system that is SN but not WN (WN implies excluded middle). 
\end{todolist}

\section*{Plan for PLDI submission}
\begin{todolist}
\item[X] Aim for first draft finished by the end of the 29th.
  \item Make a main file at this point to conform to this conference's style. Possibly a sub-folder for each conference.
  \item Submission deadline is Thursday 13th 
  \item Once we've got a current draft finished. Look in Teresa for first order term rewrite systems and see how many we can represent using our code.
  \item [\done] discussion env has qed boxes which shouldn't be there.
  \item Work on the conclusion
  \item from wellfounded extra move the implication of fb -> isdec -> isMinDec. move to implications. 
  Use this to take this SNdec→WN to have a full proof from fb, isdec, sn -> wn. (issue in moving to fb file due to reversal of relation direction!)
  \item Make consistent how we display definitions. 
  \item Final read through (wed?)
\end{todolist}

>>>>

    \item [\done] (Review 3) Check out and include the following related work:

@@

\begin{itemize}
  \item[\done] Perhaps we should not create variations and multiple implications of WFind,
  but rather prove that it's equivalent to WFacc, but lands in a higher universe.
  Mention this once at the beginning of section 5 then proceed to analyze
  connections between WFacc and WFseq/WFmin.
\end{itemize}

\section*{July TODO}
\begin{todolist}
\item [\done] Carry on refactoring Wellfounded.agda once fixed, breaking it up if necessary.
\item [\done] To finish the paper: We should do a final walk through and make sure we haven't introduced errors with our edits. In particular, do all the counterexamples line up. Do all the references to different sections and pieces of code in Agda line up. 
  \item [\done] Do we remove all our local counterexamples in favour of global ones?
  \item [\done] Rework well-foundedness section 
  \item[\done]  Rework conclusion 
  \item [\done]  Rework introduction 
  \item  [\done] Do we introduce a new section for discussing Agda? Talking about why we think there is an advantage to doing this work in Agda vs (say) Isabelle

@@

\section*{Research items}
\begin{todolist}
\item[\done] (Research item) Go deeper in the related works and see which are most relevant. Make a list of works and give summaries to be used in future papers. * Look into the literature and related works. There should be an Agda formulation of rewrite systems (check agda standard library stuff on relations). 
  \item (Research item) Find a concrete application of generalized Newman's Lemma. * Finding a term rewrite system where it's eassier to see that it is weakly cr and that every term is SM (as opposed to being CR).  
  \item (Research item) Formalize a first order term rewriting system (I can't remember what the motivation for this was). Use TeReSe to create a list of rewrite systems.
  \item (Research item) Carry out research on rewriting counterexamples like our own so we can try and put them into some sort of context. Have our counerexamples been shown before?
  \item (Research item) Work on the missing implicaitons from the paper. So far we believe this is just CR $\to$ CP in the ARS section. - Will probably be quite a bit of work (low priority)
  \item (Research item) Do we want to reincorporate Knaster Tarski? - Not this paper 
  \item (Research item) How to go from WFseq to WFacc - possible that the Kleene Tree is a solution. - Not this paper
  \item (Research item) Investigate: what we've done for natural numbers could be done for an arbitrary ARS (in the counter examples files), but for weak minimality and weak excluded middle. 
  \end{todolist}

\section*{Agda standard library}
Some of the work we have explored is formalized in the Agda standard library.  
% https://agda.github.io/agda-stdlib/v1.7.3/Relation.Binary.Rewriting.html#2096
\begin{todolist}
\item[\done] Write about the standard library formalization in our paper and talk about how ours differs. 
\item[X]  Create an equivalences file showing that our formalization agrees with the ASL (specifically definitions such as confluet, wcr, nf, leading up to showing our formalizations of Newman's lemma agree).
  \item Investigate contributing work to the ASL. This may mean refactoring the work already done to conform with the ASL. Specific contributions would likely be: Generalized NL. Theorem's 122 and 123. Alternative definitions of Wellfounded, the implications for the hierarchy.
\end{todolist}

\section*{20th August todos}
\begin{todolist}
  \item [\done] (Secondary) Find consistency with NP and UN with respect to reduction and conversion. Should the default be by reduction for both. Should we be explicit with $\to$ and = in each case.  \sacomment{Main issue here is that we'll either introduce an inconsistency with our code or we'll have a hard to read symbol. We can do $\to$ as super or sub-script in Agda. If we're happy with $NP\to$ in agda but using a different script in the paper then I think we should go ahead and be explicit for both UN and NP as to whether it is WRT conversion or reduction. We already have this inconsistency in the code with $UN\to$} \apcomment{In the agda code let's not use arrow and have arrow implied as default but be explicit with '=' sign.}
  \item [\done] Carry on modifying the names of wfXXX- to wfXXXnotnot with prefix underscore in agda and in latex. \sacomment{Done with the exception of WFseq-. This needs a bit more thought. Similarly for the relation between WFcor and WFminCor.}
  \item [\done] Make a decision regarding WFcor. We use it to imply WFacc using AccCor. But otherwise it is sort of orphaned off. Does it help contribute to the overall story? \sacomment{We've now related it to WFmin by requiring that correductive predicates have double negation elimination. Is this too strong a requirement? Similarly we have linked WFacc to WFcor but this proof requires the same property as well as limiting ourselves to accessible predicates. Again, is this too much?}
  \item [\done] Get the classical hypothesis and relations organized and see how they're used in the implications (organize implications by classical requirement. In the implications file we can import the classical implications. Basically, refactor the code.) \sacomment{Done but could be further refactored.}
  \item [\done] Make a cheat sheet for the next meeting regarding to talk about coreductive. How it relates on the weak side and on the strong side. SEQ seq+- , how to present these. Try and clarify what the issues are here and what exactly are the differences between them in terms of definition and implication. Could one be removed and the overall story stay the same? \sacomment{Regarding seq+ (and seq+-) we currently only show that seq+ implies seq and seq+- implies seq- so could be removed without changing any of our other implictions.} 
\end{todolist}

\section*{September 15th todo}
\begin{todolist}
\item[\done] (Minor) We want to have a sentence in our WF section making clear that we can consider local properties as well and maybe make an example from accessible to sequential. 
  \item[\done] Sentence explaining that wfacc is equivalent to wfind and therefore isn't included in the diagram (wfind).
\end{todolist}

\section*{September 26th TODO}
\begin{todolist}     
\item[\done] Focus for next meeting on writing. 
  \item [\done] Are not not closed predicates over natural numbers dedcidable: If not is this a counterexample for WFminDNE.
\end{todolist}

\section*{October 10th TODO}
\begin{todolist}     
\item[\done] Want to show that FB and weak decidability gives us accCor (think we may have shown this previously)
  \item[\done] Add 'is' to wfcor macros.
  \item[\done] make a decision on wfseq not not name. 
  \item[\done] Keep on cleaning the WF section in paper. 
\end{todolist}

\section*{October 15th TODO / Oct 22nd}
\begin{todolist}     
  \item It's possible to have a term rewrite system with good properties like FB and the relation being decidable. But not satisfying accCor. That is something that can be added in a final touch of the paper. Either a remark or example, where a term rewrite system does not validate accCor. (In the lambda calculus, there are computable perpetual strategies. We think they're concerned with weak normalization. Not sure if it can be done with strong normalization.)
  \item [\done] Do a final read through of the information below the figure. Have we captured all the most important information / Are we repeating ourselves?
  \item [\done] Do we want to keep anything in the old/delete section of WF?
  \item [\done] Do we want to add anything about counter examples or positive examples? 
  \item [\done] We may want to use words rather than macros in certain locations
  \item [\done] Check through the comments in general in the section. 
\item [\done]  CHange WFseq not not to WFseq- in the macro. We are sticking with this for now. If we have time we can investigate the relation between the two.
\item [\done]  Make the diagram horizontal rather than vertical. Spoof it up and we can compare next week.
\item [\done]  Use of macros in WF section. Double check the right one is being used. Don't use the "is'' macro at all in this section. Be consistent on whether we are using macros. 
\item Does WFseq- with MP== imply WFseq? (W're thinking it will need rdec as well to give us the implication).
\item [\done] make the WF macro bold font?
\item add wfseq not not to the definitions?
\item [\done] Go through the bibliography links. Not all of them are currently working!
\item We want to start looking at examples. What we want to know is whether a first order rewrite system with some minimal properties implies some of our classical properties. E.g. lambda calculus satisfies some of our classical properties. As does first order term rewrite system. Follow up question: how big of an issue is it that accCor is not provable? Can we prove is isn't satisfied? 
\item [\done] Make a connection in the final WF section talking about how SN doesn't imply WN constructively. (we've started writing this sentence). 
\item [\done] We may need to add to the WFCounters file, if we have SN implies WN then every subset of the natural numbers is decidable.  
\item [\done] Make sure we have a proof of the implication from SN to WN given FB and decidability of a relation (to back up our claim in our introduction).
\item check TERESE and see first order term rewrite systems example which could be valid for showing classical and constructive formalization coincide. First order term rewrite systems are (always?) fb and R is dec. 
\item [\done] Change the macro for FB so that it is also boldface
\item [\done, did SN to WN instead] update wfcounters to show that asserting wfmin implies the excluded middle (and wfmindne the weak excluded middle).
\item [\done] In wfcounters we should have a system that is SN but not WN (WN implies excluded middle). 
\end{todolist}

\section*{Plan for PLDI submission}
\begin{todolist}
\item[X] Aim for first draft finished by the end of the 29th.
\item[\done] Make a main file at this point to conform to this conference's style. Possibly a sub-folder for each conference.
\item [\done] discussion env has qed boxes which shouldn't be there.
\item[\done] from wellfounded extra move the implication of fb -> isdec -> isMinDec. move to implications. 
  Use this to take this SNdec→WN to have a full proof from fb, isdec, sn -> wn. (issue in moving to fb file due to reversal of relation direction!) 
\item Make consistent how we display definitions. 
\item [\done] Work on the conclusion
\item Need to fix how we put urls in for bibliography. Currently not quite working. 
\item Run through and compile everything, make sure we didn't introduce any bugs
\item Submit 
\item Add a paragraph or two giving a short overview/summary of Agda.
\item Go through the code, make naming consistent and self-documenting
\item Once we've got a current draft finished. Look in Teresa for first order term rewrite systems and see how many we can represent using our code.
\item Consider adding some more motivations 
\item Look at the seq-minus vs. seq-not-not issue 
\item Final read through (wed?)
\item Submission deadline is Thursday 13th 
\end{todolist}

### 2025-11-10 Relations/Paper/Well-foundedness.tex
Preparing to submit the paper

An abstract reduction $\to_R$ is strongly normalizing
if and only if the converse relation is well-founded.
In this section, we will compare several definitions of well-foundedness.
%which depend on the basic notions given in Definition~\ref{def:WFnotions} below.

@@

Such distinctions have consequences for rewriting theory.
For example, recall that $x\in \SN_R$ iff $x$ is $\bar{R}$-accessible, where $\bar{R}$
is the converse of $R$.  Thus, if $\bar{R}$ is well-founded, $R \models \gSN$.
Yet actually computing the normal form of a given $x$ seems to require a
strong form of decidability of being $R$-minimal, namely:
\[
\tag{$\isMinDec$} \forall x. \left(\exists y. Ryx\right) \lor \left(\forall y. \lnot Ryx\right)
\]
This is stronger than deciding whether $x$ is a normal form:
% Note that deciding whether $x$ is a normal form constitutes the weaker disjunction
$\lnot\lnot\left(\exists y. Ryx\right) \lor \left(\forall y. \lnot Ryx\right)$.
Indeed, in the file \texttt{ARS/Examples.agda} the implication $\gSN \to \gWN$ 
is equivalent to the principle of excluded middle.

>>>>

\newcommand{\gWF}{\mathbf{WF}}
\newcommand{\WF}{\mathrm{WF}}

An abstract reduction $\to_R$ is strongly normalizing
if and only if the converse relation is well-founded.
In this section, we will compare several definitions of well-foundedness.
For ease of exposition, we will only consider local versions of these notions.
(There is no loss of generality in this; any global well-foundedness property $\gWF(A;R)$ 
restricts to a predicate $WF_R$ on $A$ via $WF_R(x) = \gWF(\Sigma y : A. R^* x y;R')$, 
where $R'((y,\rho),(z,\sigma)) = R(y,z)$.)
%which depend on the basic notions given in Definition~\ref{def:WFnotions} below.

@@

Such distinctions have consequences for rewriting theory.
For example, recall that $x\in \SN_R$ iff $x$ is $\bar{R}$-accessible, where $\bar{R}$
is the converse of $R$.  Thus, if $\bar{R}$ is well-founded, $R \models \gSN$.
Yet actually computing the normal form of a given $x$ requires a
strong form of decidability of being $R$-minimal:
\[
\tag{$\isMinDec$} \forall x. \left(\exists y. Ryx\right) \lor \left(\forall y. \lnot Ryx\right)
\]
Note that merely deciding whether $x$ is a normal form is the weaker statement
% Note that deciding whether $x$ is a normal form constitutes the weaker disjunction
$\lnot\lnot\left(\exists y. Ryx\right) \lor \left(\forall y. \lnot Ryx\right)$.
Indeed, in the file \texttt{ARS/Examples.agda} the implication $\gSN \to \gWN$ for all relations $R$
is shown to be equivalent to the principle of excluded middle.
